Naive Bayes Workflow Ver 2.0.r

### 2019-02-21 docs/hiding/Multinomial_bayes_V1.Rmd
update

---
title: "Naive Bayes Workflow Ver 2.0.r"
author: "Dan Johnson"
date: "1/29/2019"
output: html_document
---

@@

##Summary
This is **Version 2.0.r** of the basic workflow for classifying TA comments using a naive Bayes model. It pre-processes text using the same command set as Ver 2.0.b. The "r" designates this version has been modified to be able to randomize which rows are selected for training and test datasets.

@@

```{r}
library(tidyverse)
library(tidytext)
library(tidyr)
library(tm)
library(SnowballC)
library(wordcloud)
library(e1071)
library(gmodels)
```

@@

##Version 2.0.r of Naive Bayes (NB) Protocol
###Initial Data Input
```{r}
#Read in TA comments from CSV file.
base_dataVERB <- read_csv(file='data/coded_full_comments_dataset_Spring18anonVERB.csv')
```

>>>>

---
title: "Multinomial Naive Bayes Workflow Ver 1.0."
author: "Dan Johnson"
date: "2/13/2019"
output: html_document
---

@@

##Summary
This is **Version 1.0** of a basic workflow for classifying TA comments using a multinomial naive Bayes model. It pre-processes text using the same command set as NB-Ver 2.0. This version replaces categorical naive Bayes with a multinomial classifier.

@@

```{r}

library(tm)
library(RTextTools)
library(e1071)
library(dplyr)
library(caret)
library(doMC)
registerDoMC(core=detectCores())

#library(tidyverse)
#library(tidytext)
#library(tidyr)
#library(SnowballC)
#library(wordcloud)
#library(gmodels)
```

@@

##Version 2.0.r of Naive Bayes (NB) Protocol
###Initial Data Input
```{r}
#Read in TA comments from CSV file.
subset_M <- read.csv("/Users/danjohnson/Dropbox/Coding_Tools/R_Environment/R_Projects/ta_comments_FLC/comments_subset.csv", stringsAsFactors=FALSE)
glimpse(subset_M)
```
```{r}
set.seed(1)
subset_M <- subset_M[sample(nrow(subset_M)), ]
subset_M <- subset_M[sample(nrow(subset_M)), ]
glimpse(subset_M)
```

```{r}
subset_M$subject <- as.factor(subset_M$subject)
subset_M$structure <- as.factor(subset_M$structure)
```

```{r}
corpus <- Corpus(VectorSource(subset_M$text))
corpus
inspect(corpus[1:3])
```

```{r}
corpus.clean <- corpus %>%
  tm_map(content_transformer(tolower)) %>% 
  tm_map(removePunctuation) %>%
  tm_map(removeNumbers) %>%
  tm_map(removeWords, stopwords()) %>%
  tm_map(stripWhitespace)
```

```{r}
dtm <- DocumentTermMatrix(corpus.clean)
# Inspect the dtm
inspect(dtm[40:50, 10:15])
```

```{r}
subset_M.train <- subset_M[1:7005,]
subset_M.test <- subset_M[7006:9340,]

dtm.train <- dtm[1:7005,]
dtm.test <- dtm[7006:9340,]

corpus.clean.train <- corpus.clean[1:7005]
corpus.clean.test <- corpus.clean[7006:9340]
```

```{r}
dim(dtm.train)
```

```{r}
fivefreq<- findFreqTerms(dtm.train, 5)
length((fivefreq))

dtm.train.nb <-DocumentTermMatrix(corpus.clean.train, control=list(dictionary=fivefreq))

dim(dtm.train.nb)

dtm.test.nb <- DocumentTermMatrix(corpus.clean.test, control=list(dictionary = fivefreq))

dim(dtm.train.nb)
```

We use a variation of the multinomial Naive Bayes algorithm known as binarized (boolean feature) Naive Bayes due to Dan Jurafsky. In this method, the term frequencies are replaced by Boolean presence/absence features. The logic behind this being that for sentiment classification, word occurrence matters more than word frequency.
```{r}
# Function to convert the word frequencies to yes (presence) and no (absence) labels
convert_count <- function(x) {
  y <- ifelse(x > 0, 1,0)
  y <- factor(y, levels=c(0,1), labels=c("No", "Yes"))
  y
}
```

```{r}
# Function to avoid converting word frequencies to yes (presence) and no (absence) labels
convert_count <- function(x) {
  y <- x
}
```

# Apply the convert_count function to get final training and testing DTMs
```{r}
trainNB <- apply(dtm.train.nb, 2, convert_count)
testNB <- apply(dtm.test.nb, 2, convert_count)
```

```{r}
# Train the classifier
system.time( classifier <- naiveBayes(trainNB, subset_M.train$subject, laplace = 1) )
#system.time( classifier <- naiveBayes(trainNB, subset_M.train$structure, laplace = 1) )
```

Testing the Predictions
```{r}
# Use the NB classifier we built to make predictions on the test set.
system.time( pred <- predict(classifier, newdata=testNB) )
```

```{r}
# Create a truth table by tabulating the predicted class labels with the actual class labels 
table("Predictions"= pred,  "Actual" = subset_M.test$subject )
```

```{r}
# Prepare the confusion matrix
conf.mat <- confusionMatrix(pred, subset_M.test$subject)
conf.mat
print ("   ")

conf.mat$byClass
print ("   ")

conf.mat$overall
print ("   ")

conf.mat$overall['Accuracy']

https://rpubs.com/cen0te/naivebayes-sentimentpolarity

### 2019-02-21 docs/hiding/Multinomial_bayes_V3_1.Rmd
update

subset_M <- read.csv("/Users/danjohnson/Dropbox/Coding_Tools/R_Environment/R_Projects/ta_comments_FLC/comments_binary.csv", stringsAsFactors=FALSE)
glimpse(subset_M)
```

```{r}
set.seed(2012) #Randomizer
subset_M <- subset_M[sample(nrow(subset_M)), ]
glimpse(subset_M)
```

@@

msg.dfm <- dfm(text.corpus, tolower = TRUE)  #generating document freq matrix
msg.dfm <- dfm_trim(msg.dfm, min_termfreqError = 5, min_docfreq = 3)  
msg.dfm <- dfm_weight(msg.dfm)

@@

@@

Training the Naive Bayes classifier-

nb.classifier<-textmodel_NB(msg.dfm.train,spam.train[,1])
nb.classifier
## Fitted Naive Bayes model:
## Call:
##  textmodel_NB.dfm(x = msg.dfm.train, y = spam.train[, 1])
## 
## 
## Training classes and priors:
## spam  ham 
##  0.5  0.5 
## 
##        Likelihoods:      Class Posteriors:
## 30 x 4 Matrix of class "dgeMatrix"
##                 spam          ham       spam       ham
## you     5.001507e-03 0.0096798156 0.34067144 0.6593286
## have    4.322289e-03 0.0042303673 0.50537386 0.4946261
## 1       2.695748e-03 0.0009529526 0.73882413 0.2611759
## new     3.492485e-03 0.0010753934 0.76457487 0.2354251
## .       6.965338e-03 0.0168302131 0.29271598 0.7072840
## please  2.339097e-03 0.0011593603 0.66860811 0.3313919
## call    1.058603e-02 0.0021859571 0.82884759 0.1711524
## i       8.439760e-04 0.0112106647 0.07001254 0.9299875
## wait    1.860817e-04 0.0011538316 0.13887596 0.8611240
## for     5.699340e-03 0.0045025239 0.55865674 0.4413433
## hope    2.334040e-04 0.0017258550 0.11912872 0.8808713
## tonight 1.137075e-04 0.0011106417 0.09287182 0.9071282
## too     3.802754e-05 0.0017024748 0.02184860 0.9781514
## bad     1.232420e-04 0.0006045270 0.16934219 0.8306578
## as      1.339518e-03 0.0020699791 0.39287852 0.6071215
## well    2.938089e-04 0.0017334850 0.14492664 0.8550734
## but     2.528948e-04 0.0043933716 0.05442968 0.9455703
## rock    3.802754e-05 0.0002684845 0.12406542 0.8759346
## night   3.003905e-04 0.0017976398 0.14317739 0.8568226
## anyway  3.802754e-05 0.0005405216 0.06572915 0.9342709
## going   1.538819e-04 0.0023951976 0.06036762 0.9396324
## a       7.856726e-03 0.0064918622 0.54756091 0.4524391
## now     6.254232e-03 0.0028758075 0.68501697 0.3149830
## good    6.723203e-04 0.0030342352 0.18138681 0.8186132
## speak   8.003838e-04 0.0004728416 0.62862694 0.3713731
## to      1.113210e-02 0.0075761991 0.59503541 0.4049646
## soon    2.642059e-04 0.0010608467 0.19939274 0.8006073
## today   1.120666e-03 0.0019041688 0.37048833 0.6295117
## is      4.451802e-03 0.0058153446 0.43359683 0.5664032
## accept  3.802754e-05 0.0003188419 0.10655871 0.8934413

The model outputs the Probabilities of the message being Spam or ham.
Let’s Test the Model

pred<-predict(nb.classifier,msg.dfm.test)

# use pred$nb.predicted to extract the class labels
table(predicted=pred$nb.predicted,actual=spam.test[,1])
##       actual
## predicted ham spam
##      ham  952    7
##      spam  16  140

@@

#acccuracy of the classifier on Test data
mean(pred$nb.predicted==spam.test[,1])*100
## [1] 97.93722

>>>>

subset_M <- read.csv("/Users/danjohnson/Dropbox/Coding_Tools/R_Environment/R_Projects/ta_comments_FLC/comments_binary.csv", stringsAsFactors=FALSE)
glimpse(subset_M)
```
```{r}
set.seed(2012) #Randomizer
subset_M <- subset_M[sample(nrow(subset_M)), ]
glimpse(subset_M)
```

@@

msg.dfm <- dfm(text.corpus, tolower = TRUE)  #generating document freq matrix
msg.dfm <- dfm_trim(msg.dfm, min_termfreqError = 5, min_docfreq = 1)  
msg.dfm <- dfm_weight(msg.dfm)

@@

head(msg.dfm.train)
head(msg.dfm.test)
```

```{r}
#library(quanteda)
nb.classifier<-textmodel_nb(msg.dfm.train, text.train[,1])
nb.classifier
```

```{r}
predGroup <- predict(nb.classifier,msg.dfm.test)
```

# use pred$nb.predicted to extract the class labels
```{r}
table(predicted=predGroup[["names"]],actual=text.test[,1])
mean(predGroup$names==text.test[,1])*100

@@

##       actual
## predicted ham spam
##      ham  952    7
##      spam  16  140

@@

#acccuracy of the classifier on Test data
## [1] 97.93722

### 2019-02-21 docs/hiding/Naive_bayes_V2r_2a.Rmd
update

---
title: "Naive Bayes Workflow Ver 2.0.r"
author: "Dan Johnson"
date: "1/29/2019"
output: html_document
---

@@

##Summary
This is **Version 2.0.r** of the basic workflow for classifying TA comments using a naive Bayes model. It pre-processes text using the same command set as Ver 2.0.b. The "r" designates this version has been modified to be able to randomize which rows are selected for training and test datasets.

@@

```{r}
library(tidyverse)
library(tidytext)
library(tidyr)
library(tm)
library(SnowballC)
library(wordcloud)
library(e1071)
library(gmodels)
```

@@

```{r}
#Generate a vector of ~10,000 numbers (be sure it can be generated consistently)
#Adjust the number so it matches the number of rows in the original data frame
#Next, set the seed for random numbers, and randomize the order of the vector
vector <- 1:9340
set.seed(123) #change this seed value, and the data are randomized in a new sequence.
vector <- sample(vector)
glimpse(vector)
```

@@

#Remove punctuation using the "removePunctuation" function. 
#This step removes evidence of questions, so may remove data.
comments_corpus_cleanVERB_R <- tm_map(comments_corpus_cleanVERB_R, removePunctuation)

@@

This code uses the comments in the order they are listed in the original file. Block for randomizing them was done earlier.
```{r}
comments_dtm_train <- comments_dtm_1gram[1:7005, ]
comments_dtm_test <- comments_dtm_1gram[7006:9340, ]
```

Save vectors labeling rows in the training and testing vectors
```{r}
comments_train_labels_subject <- comments_rawVERB_R[1:7005, ]$subject
comments_test_labels_subject <- comments_rawVERB_R[7006:9340,]$subject
```

```{r}
comments_train_labels_structure <- comments_rawVERB_R[1:7005, ]$structure
comments_test_labels_structure <- comments_rawVERB_R[7006:9340,]$structure
```

Make sure that the proportion of each sub-category is similar in the training and testing data sets.
```{r}
prop.table(table(comments_train_labels_subject))
prop.table(table(comments_test_labels_subject))
```

```{r}
prop.table(table(comments_train_labels_structure))
prop.table(table(comments_test_labels_structure))
```

@@

Limit Document Term Matrix to only include words in the comments_freq_words vector. Using all the rows, but we want to limit the columns to these words in the frequency vector.
```{r}
comments_dtm_freq_train <- comments_dtm_train[ , comments_freq_words]
comments_dtm_freq_test <- comments_dtm_test[ , comments_freq_words]
```

The basic e1071 naive bayes classifier works with categorical features, so the matrix must be converted "yes" and "no" categorical variables. This is done using a **convert_counts function** and applying it to the data. This replaces values greater than 0 with yes, and values not greater than 0 with no. 
```{r}
convert_counts2 <- function(x) {
  x <- ifelse(x > 0, "Yes", "No")
}

comments_train <- apply(comments_dtm_freq_train, MARGIN = 2, convert_counts2)
comments_test <- apply(comments_dtm_freq_test, MARGIN = 2, convert_counts2)
```

@@

@@

###Train Model, Predict, Evaluate for Subject
This block uses the e1071 package to implement Naive Bayes algorithm on the data, and predict whether a message is likely to be in group 1_basic, 2_writing, 3_technical, or 4_logic. It evaluates the prediction with the actual data using a crosstable from the gmodels package.
```{r}
comments_classifier_subject <- naiveBayes(comments_train, comments_train_labels_subject)
comments_test_pred_subject <- predict(comments_classifier_subject, comments_test)
CrossTable(comments_test_pred_subject, comments_test_labels_subject, prop.chisq = FALSE, prop.t = FALSE, dnn = c('predicted', 'actual'))
```

The cross table lists how many of the predicted categories are in their correct categories. Adding diagonally and dividing by total count gives the percent accuracy at predicting subject.
```{r}
print("Overall subject accuracy")
(43+473+1131+184)/23.35  #Need to find a way to calculate this automatically. 
```

@@

```{r}
comments_classifier_structure <- naiveBayes(comments_train, comments_train_labels_structure)
comments_test_pred_structure <- predict(comments_classifier_structure, comments_test)
CrossTable(comments_test_pred_structure, comments_test_labels_structure, prop.chisq = FALSE, prop.t = FALSE, dnn = c('predicted', 'actual'))
```

Again, a cross table shows how many of the predicted categories match the correct categories for structure.

```{r}
print("Overall structure accuracy")
(22+488+292+518+33+10)/23.35  #Need to find a way to calculate this automatically. 
```

###OPTIONAL Output
To make a table showing predicted categories for comments based on the Naive Bayes test, versus the actual label entered by a human observer, merge the two datasets into a single data frame. The full data frame will be too long to display; use "head" to check format, then output the full dataset to CSV for further review. By default the CSV file should be saved to the current working directory.

@@

```{r}
#Creates a data frame that compares predicted and actual labels for subject and structure
tally <- data.frame("subject_label"=c(comments_test_labels_subject), "subject_predicted"=c(comments_test_pred_subject), "structure_label"=c(comments_test_labels_structure), "structure_predicted"=c(comments_test_pred_structure))
#tally <- as.numeric(as.character(tally))
```

>>>>

---
title: "Naive Bayes Workflow Ver 2.0.r"
author: "Dan Johnson"
date: "2/14/2019"
output: html_document
---

@@

##Summary
This is **Version 2r.2a** of the basic workflow for classifying TA comments using a naive Bayes model. It pre-processes text using the same command set as Ver 2.0.b. The "r" designates this version has been modified to be able to randomize which rows are selected for training and test datasets.

This updated version refines the final output tables so they are more informative.

@@

```{r}
library(tidyverse)
library(tm)
library(tidytext)
library(e1071)

#library(RTextTools)
#library(caret)
#library(doMC)
#registerDoMC(core=detectCores())
#library(tidyr)
#library(dplyr)
#library(gmodels)
```

@@

```{r}
#Generate a vector of ~10,000 numbers (be sure it can be generated consistently)
#Adjust the number so it matches the number of rows in the original data frame
#Next, set the seed for random numbers, and randomize the order of the vector
vector <- 1:9340
set.seed(567) #change this seed value, and the data are randomized in a new sequence.
vector <- sample(vector)
glimpse(vector)
```

@@

#Remove punctuation using the "removePunctuation" function. 
#This step removes evidence of questions, so may remove data.
#THIS ALSO REMOVES EVIDENCE OF POINTER ITEMS
comments_corpus_cleanVERB_R <- tm_map(comments_corpus_cleanVERB_R, removePunctuation)

@@

This code uses the comments in the order they are listed in the original file. Block for randomizing them was done earlier.
```{r}
comments_dtm_train <- comments_dtm_1gram[1:7005, ]
comments_dtm_test <- comments_dtm_1gram[7006:9340, ]
comments_dtm_all <- comments_dtm_1gram[1:9340, ]
```

Save vectors labeling rows in the training and testing vectors
```{r}
comments_train_labels_subject <- comments_rawVERB_R[1:7005, ]$subject
comments_test_labels_subject <- comments_rawVERB_R[7006:9340,]$subject
comments_all_labels_subject <- comments_rawVERB_R[1:9340,]$subject
```

```{r}
comments_train_labels_structure <- comments_rawVERB_R[1:7005, ]$structure
comments_test_labels_structure <- comments_rawVERB_R[7006:9340,]$structure
comments_all_labels_structure <- comments_rawVERB_R[1:9340,]$structure
```

Make sure that the proportion of each sub-category is similar in the training and testing data sets.
```{r}
prop.table(table(comments_train_labels_subject))
prop.table(table(comments_test_labels_subject))
prop.table(table(comments_all_labels_subject))
```

```{r}
prop.table(table(comments_train_labels_structure))
prop.table(table(comments_test_labels_structure))
prop.table(table(comments_all_labels_structure))
```

@@

Limit Document Term Matrix to only include words in the comments_freq_words vector. Using all the rows, but we want to limit the columns to these words in the frequency vector.
```{r}
comments_dtm_freq_train <- comments_dtm_train[ , comments_freq_words]
comments_dtm_freq_test <- comments_dtm_test[ , comments_freq_words]
comments_dtm_freq_all <- comments_dtm_all[ , comments_freq_words]
```

The basic e1071 naive bayes classifier works with categorical features, so the matrix must be converted "yes" and "no" categorical variables. This is done using a **convert_counts function** and applying it to the data. This replaces values greater than 0 with yes, and values not greater than 0 with no. 
```{r}
convert_counts2 <- function(x) {
  x <- ifelse(x > 0, "Yes", "No")
}
```

@@

```{}
# Function to KEEP word frequencies while still using the same variable names and code structure. The resulting matrices have cells with word frequencies instead.
convert_counts2 <- function(x) {
  y <- x
}
```

```{r}
comments_train <- apply(comments_dtm_freq_train, MARGIN = 2, convert_counts2)
comments_test <- apply(comments_dtm_freq_test, MARGIN = 2, convert_counts2)
comments_all <- apply(comments_dtm_freq_all, MARGIN = 2, convert_counts2)
```

@@

###Train Model, Predict, Evaluate for Subject
This block uses the e1071 package to implement Naive Bayes algorithm on the data, and predict whether a message is likely to be in group 1_basic, 2_writing, 3_technical, or 4_logic. It evaluates the prediction with the actual data using a crosstable from the gmodels package.
```{r}
comments_classifier_subject <- naiveBayes(comments_train, comments_train_labels_subject, laplace=1)
comments_test_pred_subject <- predict(comments_classifier_subject, comments_test)
comments_all_pred_subject <- predict(comments_classifier_subject, comments_all)

# Create a truth table by tabulating the predicted class labels with the actual class labels 
table("Predictions"= comments_test_pred_subject,  "Actual" = comments_test_labels_subject )

table("Predictions"= comments_all_pred_subject,  "Actual" = comments_all_labels_subject )

#table("Predictions2"= comments_test_pred_structure,  "Actual" = comments_test_labels_structure )

#The truth table lists how many of the predicted categories are in their correct categories. Adding diagonally and dividing by total count gives the percent accuracy at predicting subject. this is what the confusion matrix calculates. The faster way is to calculate a confusion matrix, which includes BOTH the truth table, and the final overall accuracy.

#The Cross Table command below (from "gmodels") generates a more detailed output table with frequency values for each individual cell.
#CrossTable(comments_test_pred_subject, comments_test_labels_subject, prop.chisq = FALSE, prop.t = FALSE, dnn = c('predicted', 'actual'))
```

```{r}
#"line" is just a spacer text for readability.
line<-("**************************************************************")

# Prepare the confusion matrix. 
conf.mat <- confusionMatrix(comments_test_pred_subject, comments_test_labels_subject)
conf.mat2 <- confusionMatrix(comments_all_pred_subject, comments_all_labels_subject)

# This print-out has both the overall and by-group accuracy values.
conf.mat2
print (line)

# This print-out has by-group accuracy values in a single column. 
conf.mat2$byClass
print (line)

# This print-out is a single line table with overall accuracy, kappa, lower and higher limits of accuracy, accuracy null value, P value for accuracy, and Mcnemar P value.
conf.mat2$overall
print (line)

# This outputs JUST the overall accuracy value. 
conf.mat2$overall['Accuracy']

# This print-out has both the overall and by-group accuracy values.
conf.mat
print (line)

# This print-out has by-group accuracy values in a single column. 
conf.mat$byClass
print (line)

# This print-out is a single line table with overall accuracy, kappa, lower and higher limits of accuracy, accuracy null value, P value for accuracy, and Mcnemar P value.
conf.mat$overall
print (line)

# This outputs JUST the overall accuracy value. 
conf.mat$overall['Accuracy']

@@

```{}
comments_classifier_structure <- naiveBayes(comments_train, comments_train_labels_structure, laplace=1)
comments_test_pred_structure <- predict(comments_classifier_structure, comments_test)
CrossTable(comments_test_pred_structure, comments_test_labels_structure, prop.chisq = FALSE, prop.t = FALSE, dnn = c('predicted', 'actual'))
```

###OPTIONAL Output
Make a table showing original comments, predicted category for comment based on the Naive Bayes test, and the actual label entered by a human observer. Export to a CSV file, with the randomizer value in the name of the file. By default the CSV file should be saved to the current working directory.

```{r}
#Creates a data frame that compares predicted and actual labels for subject and structure
tallyall <- data.frame("text"=c(comments_rawVERB_R[1:9340,]$text),"subject_label"=c(comments_all_labels_subject), "subject_predicted"=c(comments_all_pred_subject))

#Write a CSV file out with data from this run.
write.csv(tallyall, file = "SubjectsAll_Randomizer_567.csv")

```{r}
#Creates a data frame that compares predicted and actual labels for subject and structure
tally <- data.frame("text"=c(comments_rawVERB_R[7006:9340,]$text),"subject_label"=c(comments_test_labels_subject), "subject_predicted"=c(comments_test_pred_subject))

#Write a CSV file out with data from this run.
write.csv(tally, file = "Subjects_Randomizer_567.csv")

Make a table showing predicted categories for comments based on the Naive Bayes test, versus the actual label entered by a human observer, merge the two datasets into a single data frame. The full data frame will be too long to display; use "head" to check format, then output the full dataset to CSV for further review. By default the CSV file should be saved to the current working directory.

@@
